\Large Generative Dynamics of Supreme Court Citations

### 2020-03-31 Tex/paper.tex
Updated Section 4.1.1 Covariate Terms

\author{%
  Christian S. Schmid\footnote{Department of Statistics, The Pennsylvania State University, schmid@psu.edu}%
  \and Ted Hsuan Yun Chen \footnote{Department of Political Science, The Pennsylvania State University, thc126@psu.edu}%
  \and Bruce A. Desmarais \footnote{Department of Political Science, The Pennsylvania State University, bdesmarais@psu.edu}%
  \and David R. Hunter \footnote{Department of Statistics, The Pennsylvania State University, dhunter@stata.psu.edu}%
  }

@@

 \section{The Citation Temporal Exponential Random Graph Model}
We develop a methodology that can be used to jointly test for the effects of covariates on citations---as have been studied in prior research, and test for dependence effects, as we hypothesize above. To accomplish this, we extend a model that has been developed to jointly model covariate and dependence effects in longitudinal network data---the temporal exponential random graph model (TERGM) \citep[e.g.,][]{cranmer2012complex,clark2013multimember, masket2015polarization, graif2017neighborhood}. The TERGM, as it is currently implemented \citep{leifeld2017temporal,  tergm}, is designed to model longitudinal network data in which ties among a relatively stable set of nodes are modeled at multiple time points (e.g., the countries that are and are not defensively allied at time $t$, the legislators who have and have not cosponsored each others' legislation in the current session).  The longitudinal network structures for which TERGMs are currently designed are insufficient to account for the structure of citation networks.

@@

We include three sets of dummy variables that account for the issue areas of cases. In the first set of dummy variables---sender intercepts---the variable $X^{ij}$ indicates the issue area of the sending case ($i$) (e.g., for the Criminal Procedure indicator, $X_{ij} = 1$ if case $i$ is a criminal procedure case, and $0$ otherwise). The second set of dummy variables represent receiver issue area intercepts in which $X^{ij}$ indicates the issue of the receiver case. The third set of dummy variables account for the rate of citation within each issue area and between each combination of issue areas.  There is a separate dummy variable for each directed pair of issue areas. For example the A$\rightarrow$B dummy variable $X_{ij} = 1$ if case $i$ is issue area A and case $j$ is issue area B, and 0 otherwise. Note that, given the high dimension of these dummy variables, we omit one each of the sender and recipient indicators, and 28 of the 196 mixing rate indicators. Unfortunately, this leaves these indicators largely uninterpretable, but these variables are nonetheless important to absorb variation in citations that is attributable to the legal domains of two cases in a dyad. The Issue area data comes from the Supreme Court Database (SCDB) \citep{spaeth2014supreme}. We include these variables because \citet{cross2010determinants} finds that the number of citations to Supreme Court opinions depends heavily on the issue area of the case. Table \ref{issue_area_coding} reports the issue areas covered in our data.

We also include a set of covariates that account for justice-specific variation in citations. We include two types of terms at the justice level. First, we include an indicator variable that models the rate at which justices cite themselves. This effect is modeled with a single indicator variable in which $X_{ij} = 1$ if the majority opinions in cases $i$ and $j$ were written by the same justice. Second, we use two sets of dummy variables to account for differences across justices both in terms of their tendencies to include citations in the opinions they write and to be cited by other opinions. The justice sender dummy variables indicate the identity of the justice who authored the majority or plurality opinion of case $i$. The justice recipient dummy variables indicate the identity of the justice who authored the majority or plurality opinion of case $j$. This set of variables is motivated by the use of citations by \citet{kosma1998measuring} to measure the influence of individual Supreme Court justices.

@@

  \subsubsection{Dynamics in effects}

>>>>

\author{%
  Christian S. Schmid\footnote{Department of Statistics, The Pennsylvania State University, schmid@psu.edu}%
  \and Ted Hsuan Yun Chen \footnote{Department of Political Science, The Pennsylvania State University, thc126@psu.edu}%
  \and Bruce A. Desmarais \footnote{Department of Political Science, The Pennsylvania State University, bdesmarais@psu.edu}%
 % \and David R. Hunter \footnote{Department of Statistics, The Pennsylvania State University, dhunter@stata.psu.edu}%
  }

@@

 \section{The Citation Temporal Exponential Random Graph Model}
{\color{blue} This section needs to be completely revised. Model is not a TERGM anymore. - CS} 
We develop a methodology that can be used to jointly test for the effects of covariates on citations---as have been studied in prior research, and test for dependence effects, as we hypothesize above. To accomplish this, we extend a model that has been developed to jointly model covariate and dependence effects in longitudinal network data---the temporal exponential random graph model (TERGM) \citep[e.g.,][]{cranmer2012complex,clark2013multimember, masket2015polarization, graif2017neighborhood}. The TERGM, as it is currently implemented \citep{leifeld2017temporal,  tergm}, is designed to model longitudinal network data in which ties among a relatively stable set of nodes are modeled at multiple time points (e.g., the countries that are and are not defensively allied at time $t$, the legislators who have and have not cosponsored each others' legislation in the current session).  The longitudinal network structures for which TERGMs are currently designed are insufficient to account for the structure of citation networks.

@@

%We include three sets of dummy variables that account for the issue areas of cases. In the first set of dummy variables---sender intercepts---the variable $X^{ij}$ indicates the issue area of the sending case ($i$) (e.g., for the Criminal Procedure indicator, $X_{ij} = 1$ if case $i$ is a criminal procedure case, and $0$ otherwise). The second set of dummy variables represent receiver issue area intercepts in which $X^{ij}$ indicates the issue of the receiver case. The third set of dummy variables account for the rate of citation within each issue area and between each combination of issue areas.  There is a separate dummy variable for each directed pair of issue areas. For example the A$\rightarrow$B dummy variable $X_{ij} = 1$ if case $i$ is issue area A and case $j$ is issue area B, and 0 otherwise. Note that, given the high dimension of these dummy variables, we omit one each of the sender and recipient indicators, and 28 of the 196 mixing rate indicators. Unfortunately, this leaves these indicators largely uninterpretable, but these variables are nonetheless important to absorb variation in citations that is attributable to the legal domains of two cases in a dyad. The Issue area data comes from the Supreme Court Database (SCDB) \citep{spaeth2014supreme}. We include these variables because \citet{cross2010determinants} finds that the number of citations to Supreme Court opinions depends heavily on the issue area of the case. Table \ref{issue_area_coding} reports the issue areas covered in our data.

We include one covariate that accounts for the rate at which cases that fall under the same issue area cite each other. The variable $X^{ij}$ indicates whether cases $i$ and $j$ share the same issue area, e.g., $X^{ij}=1$ if cases $i$ and $j$ are classified into the same issue area, and $0$ otherwise. The Issue area data comes from the Supreme Court Database (SCDB) \citep{spaeth2014supreme}. We include these variables because \citet{cross2010determinants} finds that the number of citations to Supreme Court opinions depends heavily on the issue area of the case. Table \ref{issue_area_coding} reports the issue areas covered in our data.

%We also include a set of covariates that account for justice-specific variation in citations. We include two types of terms at the justice level. First, we include an indicator variable that models the rate at which justices cite themselves. This effect is modeled with a single indicator variable in which $X_{ij} = 1$ if the majority opinions in cases $i$ and $j$ were written by the same justice. Second, we use two sets of dummy variables to account for differences across justices both in terms of their tendencies to include citations in the opinions they write and to be cited by other opinions. The justice sender dummy variables indicate the identity of the justice who authored the majority or plurality opinion of case $i$. The justice recipient dummy variables indicate the identity of the justice who authored the majority or plurality opinion of case $j$. This set of variables is motivated by the use of citations by \citet{kosma1998measuring} to measure the influence of individual Supreme Court justices.

We also include an indicator variable that models the rate at which justices cite themselves. Similar to the same issue area variable, this effect is modeled with a single indicator variable in which $X_{ij} = 1$ if the majority opinions in cases $i$ and $j$ were written by the same justice. The data is taken from the SCDB \citep{spaeth2014supreme}. {\color{blue} Justification for including this variable? }

@@

 % \subsubsection{Dynamics in effects}

% For the current model this section is not relevant anymore
